An in-depth study of Explainability and Adversarial Robustness for RNNs

### 2019-12-05 ex3/paper/paper.tex
Add stuff for feature correlation

Surprisingly, we can confirm that adversarial samples can be found even when considering the tight constraints mentioned before. Thus we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems. The new performance score \gls{ars} precisely captures the notion of how easily an adversary can generate adversarial samples. We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process. The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario. We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low. With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.

@@

\section{Related work}

\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly. \cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible.
\cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}

@@

Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs. Since the reasons discussed above suggest that finding adversarial flows is particularly difficult in our case, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.

We implemented the following methods for generating adversarial flows.

@@

Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero. On average, one feature gets zeroed out but it is also possible that none or more than one are left out. This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer. The result is the classifier learning to deal with missing features. As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features. The results in Table~\ref{fig:featDropout} show unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance. With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port. 
\todo{Add correlation and explain why most features have no importance}

A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965. We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.

>>>>

Surprisingly, we can confirm that adversarial samples can be found even when considering the tight constraints mentioned before. Thus we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems. The new performance score \gls{ars} precisely captures the notion of how easily an adversary can generate adversarial samples.

@@

Based on the insight gained by the feature importance and explainability methods, we develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process. The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario. We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low. With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.

@@

Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs: 
\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly. \cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible. \cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}

As finding adversarial flows is particularly difficult in our case for the reasons we discussed above, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods. We implemented the several methods for generating adversarial flows.

@@

Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero. On average, one feature gets zeroed out but it is also possible that none or more than one are left out. This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer. The result is the classifier learning to deal with missing features. As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features. The results in Table~\ref{fig:featDropout} show unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance. With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port. A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965. We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.

Feature dropout also allows us to find correlated features or -- more general -- features that contain common information:

% numerator = max(float(baseline_accuracy-accuracy_for_feature_combination), 0)
% 			denominator = max(float(baseline_accuracy-accuracy_for_feature_1), 0) + max(float(baseline_accuracy-accuracy_for_feature_2), 0)
\begin{equation}
\text{score} = \frac{\text{acc}_\text{base}-\text{acc}_\text{-both\_features}}{\left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_1}\right) + \left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_2}\right)}
\end{equation}
With $\text{acc}_\text{base}$ we denote the average accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the average accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the average accuracy if both features are omitted. The higher the resulting number, the larger the information that both features share.
